An Abstraction-based Approach to Network Synthesis

### 2016-11-01 paper/abstraction/paper.tex
another pass on 2

A second limitation is inability to support network evolution. Networks evolve all the time as nodes and links are added or removed and the addresses of services are changed. A system that supports evolution should ensure that correctly supporting such changes requires only small changes to network configurations. Engineers want to minimize changes to device configurations because each change runs the risk of poor, dynamic behavior during the change.

Concrete synthesis tools do not accommodate evolution well. When engineers want to modify the network, they would use the new topology as input to the system. (Old inputs or configurations are not supplied.) In the resulting output, existing device configurations may undergo unnecessary changes, beyond those needed to accommodate the modification. Suppose we want to modify the prefix of local ToR T8 in  Figure~\ref{fig:example}. To accommodate the modification, Propane, a concrete synthesis tool, will not only change T8's configuration but also those of the two spines because it uses prefix lists at spines to ensure that local prefixes are not advertised externally.  As we show later, there is a way to design the configurations of this network in a way that does not require spine configuration changes to accommodate ToR prefix modifications, by exposing the inherent structure in this topology to the synthesis tool.

@@

While templates operate at the right level of abstraction, today's template systems have two limitations. The first is that they offer no guarantees on correctness. To author correct templates, engineers must do manually the conceptually challenging task of breaking down complex, global policies into role templates such that the collective behavior of the templates (upon compilation to configurations) correctly implements the policy under all failure conditions.

As a result, the types of configuration design errors pointed in prior work~\cite{propane} can occur in template design as well. For instance, to keep local prefixes internal the engineer may implement announcement forwarding on spines based on the identity of the neighbor, i.e., do not announce externally anything heard from local aggregators. This implementation is appealing because it does not require changes to spine configurations when ToR prefixes change. But it works correctly only when there are no failures; local prefixes can leak externally when failures disconnect a spine from both aggregators in a local pod. Specifically, when S1 is disconnected from A5 and A6, it may hear about P5 and P6 from A3 and A4 (via A[5--6]->S2->A[3--4]), and it will then announce those prefixes outside (since they no not come from a local aggregator), leading to a policy violation.

%A correct implementation here would be for each spine to not advertise externally announcements from local aggregators {\em and} reject announcements that have traversed the other spine (i.e., disallow "valleys," or up-down-up paths).

@@

The second limitation of today's template systems is that, like concrete synthesis tools, they do not support network evolution. In theory, templates are appealing because operators do not have to redesign templates when the topology changes while keeping its structure of roles and their inter-connectivity. In practice, however, there is no way to \todo{ ... continue here ...}

More importantly, because the analysis is done over a concrete topology, it provides no insight as to when the templates can be safely re-used. Templates that work for a given concrete topology may not work for seemingly-minor variations of the topology. Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}; it has the same folded-Clos structure and the same five roles. One might think that the same templates (but different database entries) can be used for both topologies. However, templates that disallow valley paths lead to a less fault-tolerant system for Figure~\ref{fig:example2}.

In particular, aggregation-induced blackholes can occur in Figure~\ref{fig:example2} with only one link failure, while it would take two failures in Figure~\ref{fig:example}. Such blackholes occur when a switch announces an aggregate prefix but has no valid path to a more specific prefix within the aggregate. In Figure~\ref{fig:example}, a blackhole will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes by virtue of announcing the encompassing aggregate PG. In contrast, in Figure~\ref{fig:example}, no such blackhole will occur after a single failure.

Thus, engineers today must reason about the correctness of templates for each concrete topology and each evolution. Worse, for some evolutions, they may discover that their old templates will not work anymore and need modifications. This discovery is highly onerous because changing the templates implies that the configuration of all the devices needs to change, which can be an unacceptable disruption.  When that happens, many networks abandon templates entirely and revert to modifying individual configuration files, using localized patches to accomodate evolution. This reversion brings backs the class of errors that the templates were designed to prevent.

As an aside, a possible template implementation that works correctly for both networks is: allow valley paths and use an explicit list of local prefixes at spines to decide which announcements should not be forwarded externally.

@@

\sysname addresses the challenges above with both template systems and concrete synthesis tools. It takes as input network policy and an {\em abstract} topology  that captures the different device roles and connectivity. It outputs a template for each role. Engineers can then use \sysname's template compiler or their own to convert templates into configurations based on the concrete topology and parameter values.

\sysname guarantees that the templates are correct, that is, when compiled to device configurations, the network will correctly implement the specified policy. These guarantees hold for any valid concrete topology that matches the specified abstract topology; and when a valid concrete topology evolves to another valid concrete topology, only necessary changes are needed to device configurations.

>>>>

A second limitation is inability to support network evolution. Networks evolve all the time as nodes and links are added or removed and the addresses of services are changed. A system that supports evolution would ensure that supporting such evolution requires only small changes to network configurations. Engineers want to minimize changes to device configurations because each change runs the risk of poor, transient behavior due to poor router software or routing protocol dynamics~\cite{ratulbgpmisconfigs}.

Concrete synthesis tools do not accommodate evolution well. When engineers want to modify the network, they would use the new topology as input to the system; old inputs or configurations are not uses. In the resulting output, existing device configurations may undergo unnecessary changes, beyond those needed to accommodate the modification. Suppose we want to modify the prefix of local ToR T8 in  Figure~\ref{fig:example}. To accommodate the modification, Propane, a concrete synthesis tool, will not only change T8's configuration but also those of the two spines because it uses prefix lists at spines to ensure that local prefixes are not advertised externally.  As we show later, there is a way to design the configurations of this network in a way that does not require spine configuration changes to accommodate ToR prefix (and other) modifications, by exposing the inherent structure in this topology to the synthesis tool.

@@

While templates operate at the right level of abstraction, today's template systems have two limitations. The first is that they offer no guarantees on correctness. To author correct templates, engineers must do manually the conceptually challenging task of breaking down complex, global policies into role templates such that the collective behavior of configurations derived from the templates correctly implements the policy under all failure conditions.

As a result, the types of configuration design errors pointed in prior work~\cite{propane} can occur in template design as well. For instance, to keep local prefixes internal the engineer may implement announcement forwarding on spines based on the identity of the neighbor, i.e., do not announce externally anything heard from local aggregators. This implementation is appealing because it does not require changes to spine configurations when ToR prefixes change. But it works correctly only when there are no failures; local prefixes can leak externally when failures disconnect a spine from both aggregators in a local pod. Specifically, when S1 is disconnected from A5 and A6, it may hear about P5 and P6 from A3 and A4 (via A[5--6]->S2->A[3--4]), and it will then announce those prefixes outside (since they no not come from a local aggregator), leading to a policy violation. A possible correct implementation here would be for each spine to not advertise externally announcements from local aggregators {\em and} reject announcements that have traversed the other spine (i.e., disallow "valleys," or up-down-up paths).

@@

The second limitation of today's template systems is that, like concrete synthesis tools, they do not support network evolution. In theory, being role-based, templates do not have to be redesigned when the network evolves while keeping its basic structure of roles and their inter-connectivity. In practice, however, it is not guaranteed that templates that work for a given topology also work for seemingly-minor variations of the topology.

Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}, with the same folded-Clos structure and five roles. One might think that the same templates (with different database entries) can be re-used across topologies. However, that is not necessarily true. For instance, templates that disallow valley paths lead to a less fault-tolerant system for Figure~\ref{fig:example2}. In particular, aggregation-induced blackholes~\cite{x} can occur in Figure~\ref{fig:example2} with only one link failure, while it would take two failures in Figure~\ref{fig:example}. Such blackholes occur when a switch announces an aggregate prefix but has no valid path to a more specific prefix within the aggregate. In Figure~\ref{fig:example}, a blackhole will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes by virtue of announcing the encompassing aggregate PG. In contrast, in Figure~\ref{fig:example}, no such blackhole will occur after a single failure.

Thus, engineers today must reason about the correctness of templates for each evolution (using expensive simulations because current systems do not provide any correctness guarantee by design). Worse, for some evolutions, they may discover that their old templates will not work anymore and need modifications. This discovery is highly problematic because changing the templates implies that the configuration of all the devices needs to change, which can be an unacceptable disruption.  When that happens, many networks abandon templates entirely and revert to hand-crafting configuration files, using localized patches to accommodate evolution. This reversion brings backs the complexity and errors that the templates were meant to prevent.

%As an aside, a possible template implementation that works correctly for both networks is: allow valley paths and use an explicit list of local prefixes at spines to decide which announcements should not be forwarded externally.

\begin{figure}
\begin{tabular}{p{0.7in}|p{0.7in}p{0.7in}p{0.7in}}
& Abstraction level & Correctness guarantees & Support evolution  \\ \hline
Concrete synthesis & \xmark & \cmark & \xmark \\ \hline
Template systems & \cmark & \xmark & \xmark \\ \hline
\sysname & \cmark & \cmark & \cmark \\ 
\end{tabular}
\caption{Properties of different methods for generating device configurations.}
\end{figure}

@@

As Table~\ref{fig:marketing} shows, \sysname addresses the limitations with both concrete synthesis and current template systems. It operates at the right level of abstraction, guarantees the correctness of configurations, and supports network evolution. It takes as input network policy and an {\em abstract} topology  that describes device roles and connectivity. It outputs a template for each role, which can be compiled to configurations based on the concrete topology and parameter values. \sysname guarantees that its templates are correct, i.e., the configurations generated from them correctly implement the specified policy for {\em any} concrete topology that matches the abstract topology; and when a valid concrete topology evolves to another valid concrete topology, only necessary changes are needed to device configurations.

The next section introduces \sysname using the example in Figure~\ref{fig:example}, and the following sections provide more detail on how we generate the templates and formally define the correctness guarantees.
